---
title: "Advertising Response Measurement"
output: flexdashboard::flex_dashboard
runtime: shiny
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE, eval = TRUE, comment="", warning = FALSE, message = FALSE, tidy.opts=list(width.cutoff=55), tidy = TRUE)
set.seed(123) # ensures repeatable results for attribution rules
options(scipen=999) # suppress scientific notation
```

```{r libraries, echo=FALSE}
library(prettydoc)
library(data.table)
library(Hmisc)
library(scales)
library(DT)
library(lubridate)
```

```{r my_functions, echo=FALSE}
fit.model <- function(dt, outcome.name, input.names, model.type, digits = 3){
  the.formula <- reduce.formula(dt = dt, outcome.name = outcome.name, input.names = input.names)
  
  if(model.type == "logistic"){
    mod <- glm(formula = the.formula, family = "binomial", data = dt)
    mod.summary <- logistic.regression.summary(glm.mod = mod, digits = digits)
  }
  if(model.type == "linear"){
    mod <- lm(formula = the.formula, data = dt)
    mod.summary <- linear.regression.summary(lm.mod = mod, digits = digits)
  }
  mod.summary.rounded <- mod.summary[, lapply(X = .SD, FUN = "round.numerics", digits = digits)]
  return(mod.summary.rounded)
}


logistic.regression.summary <- function(glm.mod, digits = 3){
  library(data.table)
  glm.coefs <- as.data.table(summary(glm.mod)$coefficients, keep.rownames = TRUE)
  alpha = 0.05
  z <- qnorm(p = 1-alpha/2, mean = 0, sd = 1)
  glm.coefs[, Odds.Ratio := exp(Estimate)]
  glm.coefs[, OR.Lower.95 := exp(Estimate - z * `Std. Error`)]
  glm.coefs[, OR.Upper.95 := exp(Estimate + z * `Std. Error`)]
  return(glm.coefs[])
}


linear.regression.summary <- function(lm.mod, digits = 3, alpha = 0.05) {
  lm.coefs <- as.data.table(summary(lm.mod)$coefficients,keep.rownames = TRUE)
  setnames(x = lm.coefs, old = "rn", new = "Variable")
  z <- qnorm(p = 1 - alpha/2, mean = 0, sd = 1)
  lm.coefs[, Coef.Lower.95 := Estimate - z * `Std. Error`]
  lm.coefs[, Coef.Upper.95 := Estimate + z * `Std. Error`]
  return(lm.coefs)
}

create.formula <- function(outcome.name, input.names, input.patterns = NA, all.data.names = NA, return.as = "character"){
  
  variable.names.from.patterns <- c()
  if(!is.na(input.patterns[1]) & !is.na(all.data.names[1])){
    pattern <- paste(input.patterns, collapse = "|")
    variable.names.from.patterns <- all.data.names[grep(pattern = pattern, x = all.data.names)]
  }
  all.input.names <- unique(c(input.names, variable.names.from.patterns))
  all.input.names <- all.input.names[all.input.names != outcome.name]
  
  if(!is.na(all.data.names[1])){
    all.input.names <- all.input.names[all.input.names %in% all.data.names]
  }

  input.names.delineated <- sprintf("`%s`", all.input.names)
  the.formula <- sprintf("`%s` ~ %s", outcome.name, paste(input.names.delineated, collapse = "+"))
  
  if(return.as == "formula"){
    return(as.formula(the.formula))
  }
  if(return.as != "formula"){
    return(the.formula)
  }
}

reduce.formula <- function(dt, outcome.name, input.names, input.patterns = NA, max.input.categories = 20, max.outcome.categories.to.search = 4, return.as = "formula"){
  require(data.table)
  dt <- setDT(dt)
  
  if(!(outcome.name %in% names(dt))){
    return("Error:  outcome.name is not in names(dt).")
  }
  
  pattern.names <- list()
  if(!is.na(input.patterns[1])){
    for(i in 1:length(input.patterns)){
      pattern.names[[i]] <- names(dt)[grep(pattern = input.patterns[i], x = names(dt))]
    }
  }
  all.input.names <- c(input.names, as.character(pattern.names))
  
  num.outcome.categories <- dt[!is.na(get(outcome.name)), length(unique(get(outcome.name)))]
  
  if(num.outcome.categories <= max.outcome.categories.to.search){
    num.unique.tab <- dt[, lapply(X = .SD, FUN = function(x){return(length(unique(x[!is.na(x)])))}), .SDcols = input.names, by = outcome.name]
    min.categories.tab <- num.unique.tab[, lapply(X = .SD, FUN = "min"), .SDcols = input.names]
    
    reduced.inputs <- names(min.categories.tab)[min.categories.tab >= 2]
  }
  if(num.outcome.categories > max.outcome.categories.to.search){
    reduced.inputs <- all.input.names
  }
  
  the.formula <- create.formula(outcome.name = outcome.name, input.names = reduced.inputs, all.data.names = names(dt), input.patterns = NA, return.as = return.as)
  return(the.formula)
}
round.numerics <- function(x, digits = 0, nearest = 1){
  if(is.numeric(x)){
    return(nearest * round(x = x/nearest, digits = digits))
  }
  else{
    return(x)
  }
}
```

```{r read_data_intro, echo=FALSE, eval=TRUE, results='hide'}
customer <- fread(input = "../Data/customer.csv")
impressions <- fread(input = "../Data/impressions.csv")
transactions <- fread(input = "../Data/transactions.csv")
```

```{r preprocessing_customer, echo=FALSE, comment=""}
customer[,past.purchase := as.factor(past.purchase)]
customer[,email := as.factor(email)]
customer[,direct := as.factor(direct)]
```

```{r preprocessing_impressions, echo=FALSE, comment=""}
impressions[,date := as.Date(date)]
impressions[,channel := as.factor(channel)]
impressions[,click := as.factor(click)]
```

```{r preprocessing_transactions, echo=FALSE, comment=""}
transactions[,V1 := NULL]
transactions[,date := as.Date(date)]
transactions[,last.touch := as.factor(last.touch)]
transactions[,last.click := as.factor(last.click)]
```


```{r constants, echo=FALSE, eval=TRUE, results='hide'}
id.name = 'id'
past.purchase.name = 'past.purchase'
email.name = 'email'
Email.name = 'Email'
Email.Holdout.name = 'Email Holdout'
direct.name = 'direct'
Direct.name= 'Direct'
Display.name = 'Display'
Social.name = 'Social'
date.name = 'date'
channel.name = 'channel'
click.name = 'click'
last.touch.name = 'last.touch'
last.click.name = 'last.click'
Day.of.Week.name = 'Day of Week'
Email.ad.effect.name = 'Email ad.effect'
Display.ad.effect.name = 'Display ad.effect'
Direct.ad.effect.name = 'Direct ad.effect'
Social.ad.effect.name = 'Social ad.effect'

unique.past.purchase.name = customer[,unique(get(past.purchase.name))]
unique.email.name = customer[,unique(get(email.name))]
unique.direct.name = customer[,unique(get(direct.name))]

test.date.list = c('2017-01-03','2017-01-17', '2017-01-24', '2017-01-31', '2017-02-07', '2017-02-14', '2017-02-21')

time.list = c('2017-01-01','2017-01-02','2017-01-03','2017-01-04','2017-01-05','2017-01-06','2017-01-07','2017-01-08','2017-01-09','2017-01-10','2017-01-11','2017-01-12','2017-01-13','2017-01-14','2017-01-15','2017-01-16','2017-01-17','2017-01-18','2017-01-19','2017-01-20','2017-01-21','2017-01-22','2017-01-23','2017-01-24','2017-01-25','2017-01-26','2017-01-27','2017-01-28','2017-01-29','2017-01-30','2017-01-31','2017-02-01','2017-02-02','2017-02-03','2017-02-04','2017-02-05','2017-02-06','2017-02-07','2017-02-08','2017-02-09','2017-02-10','2017-02-11','2017-02-12','2017-02-13','2017-02-14','2017-02-15','2017-02-16','2017-02-17','2017-02-18','2017-02-19','2017-02-20','2017-02-21','2017-02-22','2017-02-23','2017-02-24','2017-02-25','2017-02-26','2017-02-27','2017-02-28')
```

```{r Data.Prepration, echo=FALSE, eval=TRUE, results='hide'}
# Data Preparation: summarize impressions and transactions by date
trans.by.day <- xtabs(~date, transactions)
data.chanel <- xtabs(~date+channel, impressions)
mdata <- data.frame(cbind(Sales=trans.by.day[1:57], data.chanel[2:58,])) 

# Day of week
mdata$dayofweek <- weekdays(as.Date(rownames(mdata)))

# Add Ad effect variable 
mdata$Email.ad.effect <- as.numeric(filter(x=mdata$email, filter=0.5, method="recursive"))
mdata$Display.ad.effect <- as.numeric(filter(x=mdata$display, filter=0.3, method="recursive"))
mdata$Direct.ad.effect <- as.numeric(filter(x=mdata$direct, filter=0.75, method="recursive"))
mdata$Social.ad.effect <- as.numeric(filter(x=mdata$social, filter=0.3, method="recursive"))

# change variable names
colnames(mdata)<-c("Sales","Direct","Display","Email","Email Holdout","Social","Day of Week","Email ad.effect","Display ad.effect","Direct ad.effect","Social ad.effect")

# Add the weekday variable in the transactions and impressions tables so that we could inspect weekly patterns later
transactions <- transactions[, eval(Day.of.Week.name) := weekdays(as.Date(date,'%d-%m-%Y'))]
impressions <- impressions[, eval(Day.of.Week.name) := weekdays(as.Date(date,'%d-%m-%Y'))]
```

Exploratory Analysis
===================================

**Exploratory analysis upon the day of a week**

Density plots to display the distributions of different variables over the week

Row {data-height=100}
-------------------------------------

```{r EDA, echo=FALSE, eval=TRUE}
weekday_variables = c(last.touch.name,last.click.name)

inputPanel(
  selectInput(inputId = "weekday_variables", label = "Select the variable", 
              choices = weekday_variables, selected = weekday_variables[1])
)

renderPlot({
  theme_set(theme_classic())
  g <- ggplot(transactions, aes(get(Day.of.Week.name)))
  g + geom_bar(aes(fill=get(input$weekday_variables)), width = 0.5) + labs(title="Density plot", subtitle=paste(input$weekday_variables,"distribution over the week")) + theme(axis.title.x = element_blank()) + scale_fill_discrete(name = "Channels")
})
```

Last-touch for subgroups
===================================

**Last-touch analysis result for subgroups of transactions**

Take a look at who our consumers are based on whether they have past purchase record and whether they can be reached out by email or direct mail.

Row {data-height=500}
-------------------------------------

```{r Last-touch, echo=FALSE, eval=TRUE}
inputPanel(
  selectInput(inputId = "Last_touch_1", label = "Any past purchase record?", choices = unique.past.purchase.name, selected = unique.past.purchase.name[1]),
  selectInput(inputId = "Last_touch_2", label = "If email marketing is available:", choices = unique.email.name, selected = unique.email.name[1]),
  selectInput(inputId = "Last_touch_3", label = "If direct mail is available:", choices = unique.direct.name, selected = unique.direct.name[1]),
  
  checkboxInput(inputId = "Last_touch_4", label = "Sorted", value=TRUE),
  checkboxInput(inputId = "Last_touch_5", label = "Show number", value = TRUE)
)

renderPlot({
  selected.id.list = customer[get(past.purchase.name) == input$Last_touch_1 & get(email.name) == input$Last_touch_2 & get(direct.name) == input$Last_touch_3, id]
  tab <- transactions[get(id.name) %in% selected.id.list,.N,by = last.touch]
  
  if(input$Last_touch_4 == TRUE){
    setorderv(x = tab, cols = "N", order = -1)
  }  
  
  barplot <- barplot(height = tab[,N], space=0.01, las = 1, main = "Last Touch Attribution", ylab = "Transactions", xlab = "channel", ylim = c(0, 1.2*max(tab[,N], na.rm = TRUE)), col = "dodgerblue")
  text(barplot, par("usr")[3], labels = tab[,last.touch], srt = 45, adj = c(1.1,1.1), xpd = TRUE)
  
  if(input$Last_touch_5 == TRUE){
  space_val = 0
  text(x = -0.4 + 1:length(tab[,last.touch]) * (1+space_val), y = tab[,N], labels = tab[,N], pos = 3)

  }
})
```

Last-touch for different period
===================================

**Last-touch analysis result for different period of time**

Take a look at different last-touch analysis result within different period of time.

Row {data-height=500}
-------------------------------------

```{r Last-touch-1, echo=FALSE, eval=TRUE}
inputPanel(
  selectInput(inputId = "Last_touch_1_1", label = "The earliest time:", choices = time.list, selected = time.list[1]),
  selectInput(inputId = "Last_touch_1_2", label = "The ending time:", choices = time.list, selected = time.list[-1]),
  
  checkboxInput(inputId = "Last_touch_1_3", label = "Sorted", value=TRUE),
  checkboxInput(inputId = "Last_touch_1_4", label = "Show number", value = TRUE)
)

renderPlot({
  tab <- transactions[date>=input$Last_touch_1_1 &date<=input$Last_touch_1_2,.N,by = last.touch]
  
  if(input$Last_touch_1_3 == TRUE){
    setorderv(x = tab, cols = "N", order = -1)
  }  
  
  barplot <- barplot(height = tab[,N], space=0.01, las = 1, main = "Last Touch Attribution", ylab = "Transactions", xlab = "channel", ylim = c(0, 1.2*max(tab[,N], na.rm = TRUE)), col = "dodgerblue")
  text(barplot, par("usr")[3], labels = tab[,last.touch], srt = 45, adj = c(1.1,1.1), xpd = TRUE)
  
  if(input$Last_touch_1_4 == TRUE){
  space_val = 0
  text(x = -0.4 + 1:length(tab[,last.touch]) * (1+space_val), y = tab[,N], labels = tab[,N], pos = 3)

  }
})
```

Holdout Test
===================================

**Any difference about how much advertising increases sales**

The ad response over time for different email holdout test and for different duration

Row {data-height=500}
-------------------------------------

```{r hold-out test, echo=FALSE, eval=TRUE}
inputPanel(
  selectInput(inputId = "Holdout_test_1", label = "Select holdout test date:", choices = test.date.list, selected = test.date.list[1]),
  sliderInput(inputId = "Holdout_test_2", label = "Select duration period:", min = 1, max = 20, value = 7, step = 1)
)

renderPlot({
  test.date <- as.Date(input$Holdout_test_1)
  test.id <- impressions[date == test.date & channel == "email",.(id = unique(id))]
  test.id[,group := "test group"]
  hold.out.id <- impressions[date == test.date & channel == "email.holdout",.(id = unique(id))]
  hold.out.id[,group := "control group"]
  total.tab <- rbind(test.id,hold.out.id)
  total.tab[,group := as.factor(group)]
  duration <- input$Holdout_test_2
  trans.id <- transactions[date>=test.date & date<(test.date+duration),id]
  total.tab[,converted := id %in% trans.id]
  setnames(total.tab, old = "converted", new = "consumed")

  ttable <- xtabs(~group + consumed, data = total.tab)
  ttable
  mosaicplot(~group + consumed, data = total.tab,color = TRUE,
            main = paste("Holdout test on", test.date))
})
```

Marketing mix model - Basic model
===================================

**Regression relating transactions to impressions**

In order to investigate the correlations between total sales (trainsactions) and impression factors within a specific time period, we implemented linear regression model for marketing mix modeling.

Row {data-height=500}
-------------------------------------

```{r Basic Regression, echo=FALSE, eval=TRUE}
inputPanel(
  selectInput(inputId = "Basic_Regression_1", label = "Select input variables:", choices = 
c(Direct.name,Display.name,Email.name,Email.Holdout.name,Social.name,Day.of.Week.name), selected = c(Direct.name,Display.name,Email.name,Email.Holdout.name,Social.name,Day.of.Week.name), multiple = TRUE),
  sliderInput(inputId="Basic_Regression_2", label = "Select Digits:", min = 0, max = 5, value = 3, step = 1)
)

renderDataTable({
model<-fit.model(dt=mdata, outcome.name = "Sales", 
                   input.names=input$Basic_Regression_1,
                   model.type="linear", digits = input$Basic_Regression_2)
datatable(model)
})
```

Marketing mix model - Advanced model
===================================

**Account for the advertising effect and interactions between different impression factors**

Account for the decay of advertising: An ad had its biggest advertising effect just after it is shown to the users and then the effect wears over time.

Interactions occurs when there is an extra effect to have two advertising channels or impressions active meanwhile.

Row {data-height=500}
-------------------------------------

```{r Advanced Regression, echo=FALSE, eval=TRUE}
inputPanel(
  selectInput(inputId = "Advanced_Regression_1", label = "Select input variables:", choices = 
c(Email.ad.effect.name,Display.ad.effect.name,Direct.ad.effect.name,Social.ad.effect.name), selected = c(Email.ad.effect.name,Display.ad.effect.name,Direct.ad.effect.name,Social.ad.effect.name), multiple = TRUE),
  selectInput(inputId = "Advanced_Regression_2", label = "Select interaction variable 1:", choices = c(Email.ad.effect.name,Display.ad.effect.name,Direct.ad.effect.name,Social.ad.effect.name), selected = Email.ad.effect.name),
  selectInput(inputId = "Advanced_Regression_3", label = "Select interaction variable 2:", choices = c(Email.ad.effect.name,Display.ad.effect.name,Direct.ad.effect.name,Social.ad.effect.name), selected = Social.ad.effect.name),
  sliderInput(inputId="Advanced_Regression_4", label = "Select Digits:", min = 0, max = 5, value = 3, step = 1)
)

renderDataTable({
# Add Interaction
mdata$inter = mdata[,get(input$Advanced_Regression_2) * get(input$Advanced_Regression_3)]
setnames(mdata, old = 'inter', new = paste0(eval(input$Advanced_Regression_2),"*",eval(input$Advanced_Regression_3)))

model<-fit.model(dt=mdata[10:nrow(mdata),], outcome.name = "Sales", 
                   input.names=c(input$Advanced_Regression_1,paste0(eval(input$Advanced_Regression_2),"*",eval(input$Advanced_Regression_3))),
                   model.type="linear", digits = input$Advanced_Regression_4)
datatable(model)
})
```